Fail Faster

### 2025-03-21 paper/paper.tex
consistent-ify the name usage, fixing joe mistakes

In Section~\ref{section:motiv}, we introduce monadic generator DSLs, discuss how they work, and take an example-guided tour through some of the ways that
they are slower than is possible.
Then, in Section~\ref{section:faster-generators}, we explain multi-stage programming,
and develop both facets of \name{}, using \camlname{} to illustrate.
Section~\ref{section:eval} evaluates all of the previous work, demonstrating the generator performance improvements
and how they lead to much faster bug finding.

@@

\section{Evaluation}
\label{section:eval}
% \jwc{Initial section here about eval}\tr{todo}
To assess the impact of our approach, we evaluate the performance and bug-finding speed of 
\name generators across a range of benchmarks. Our experiments compare generators 
built using our technique---with and without \csplitmix---against those implemented with 
existing generator libraries (and their default randomness mechanisms). Specifically, 
we implement 
two fully semantically 
equivalent staged generator libraries that replicate the behavior of \bq 
in OCaml and ScalaCheck in Scala, allowing us to assess the effectiveness of our 
technique across different languages and runtime environments.

@@

We begin by noting that the \name{} principle of choosing fast random number generators
is completely language-agnostic. In languages like OCaml, where runtime value representations
make natively-implemented \rand inefficient, calling out to a C implementation or using standard libraries written in C is a surefire win.
In other languages, serious thought should be put into using as fast of a \rand{} as possible,
as opposed to picking up any suitable option off the shelf.

Next, we discuss language-by-language the degree to which the \name{} insights about staging could be used.
\begin{itemize}
  \item \textbf{Racket} is the next target that we intend to test on.
  The entire Racket philosophy is intertwined with
  using macros to build small eDSLs like the ones we use to write generators, and so it is a natural fit. Indeed,
  there are at least two \cn{} PBT libraries in Racket.
  However, the racket macros literature, while extensive, does not concern itself with staging for performance purposes.
  For this reason, we --- the authorship team of this paper, devoid of much Racket expertise --- decided to not use Racket as a test case in this paper.

>>>>

In Section~\ref{section:motiv}, we introduce monadic generator DSLs, discuss how they work, and take an example-guided tour through some of the ways that
they are slower than is possible.
Then, in Section~\ref{section:faster-generators}, we explain multi-stage programming,
develop \name{} (using \camlname{} to illustrate), and discuss how we isolate the performance of \rand{} to benchmark their impact on bug-finding speed.
Section~\ref{section:eval} evaluates all of the previous work, demonstrating the generator performance improvements
and how they lead to much faster bug finding.

@@

\section{Evaluation}
\label{section:eval}
% \jwc{Initial section here about eval}\tr{todo}
To assess the impact of our approach, we evaluate the performance and bug-finding speed of 
\name generators across a range of benchmarks. Our experiments compare generators 
built using our technique---with and without \csplitmix \jwc{this is the first mention of csplitmix!}---against those implemented with 
existing generator libraries (and their default randomness mechanisms). Specifically, 
we implement 
two fully semantically 
equivalent staged generator libraries that replicate the behavior of \bq 
in OCaml and ScalaCheck in Scala, allowing us to assess the effectiveness of our 
technique across different languages and runtime environments.

@@

We begin by noting that the principle of choosing fast random number generators
is completely language-agnostic. In languages like OCaml, where runtime value representations
make natively-implemented \rand inefficient, calling out to a C implementation or using standard libraries written in C is a surefire win.
In other languages, serious thought should be put into using as fast of a \rand{} as possible,
as opposed to picking up any suitable option off the shelf.

Next, we discuss language-by-language the degree to which the staging-related insights of this paper are portable.
\begin{itemize}
  \item \textbf{Racket} is the next target that we intend to test on.
  The entire Racket philosophy is intertwined with
  using macros to build small eDSLs like the ones we use to write generators, and so it is a natural fit. Indeed,
  there are at least two \cn{} PBT libraries in Racket.
  However, the racket macros literature, while extensive, does not concern itself with staging for performance purposes.
  For this reason, we --- the authorship team of this paper, devoid of much Racket expertise --- decided to not use Racket as a test case in this paper.
